The Folly of Passwords

### 2014-12-12 paper.tex
written up on OpenPGP card

\begin{itemize}
  \item OTP Apps. This is functionally equivalent to a physical OTP token,
    so a single evaluation was done on OTP apps only with the assumption that
    many aspects will generalise to any OTP generator. Some consideration
    was made around SMS OTP, but this was not evaluated directly.
  \item PKI Tokens. A smartcard-based PKI token was evaluated with the
    assumption again this will extrapolate to any PKI token. Mobile PKI was
    ruled out for consideration as it requires mobile operators to install
    keys on their customers SIM cards, which has been problematic in the
    past.
  \item PAM USB.
\end{itemize}

@@

An advantage this module has over the OATH OTP module is that
local applications can make authentication requests
multiple times without the user needing to be involved. As
long as the USB device remains connected, applications
can authenticate multiple times, but OTP authentication
will require a new password to be entered on each authentication
request.

@@

>>>>

\begin{itemize}
  \item OTP Apps. This is functionally equivalent to a physical OTP token,
    so a single evaluation was done on OTP apps only with the assumption that
    many aspects will generalise to any OTP generator. Some consideration
    was made around SMS OTP, but this was not evaluated directly.
  \item PKI Tokens. A smartcard-based PKI token was evaluated with the
    assumption again this will extrapolate to any PKI token. Mobile PKI was
    ruled out for consideration as it requires mobile operators to install
    keys on their customers SIM cards, which has been problematic in the
    past.
  \item PAM USB.
  \item YubiKey. This is a hardware USB device from Yubico that can act
    as an OTP generator, a challenge-response token and even a store
    for a static password.
\end{itemize}

@@

%An advantage this module has over the OATH OTP module is that
%local applications can make authentication requests
%multiple times without the user needing to be involved. As
%long as the USB device remains connected, applications
%can authenticate multiple times, but OTP authentication
%will require a new password to be entered on each authentication
%request.

An advantage this module has over the OATH OTP module is that
a USB storage device can also store other authentication
information such as SSH keys, SSL client certificates or
anything else to aid the Single Sign-On experience. This
increases the impact of the device being lost unless
these private keys are secured with a password, which
reintroduces passwords back into the user's workflow. Alternatively,
the risks of loss could be mitigated with an effective
key and certificate revocation process.

Futher research may find a solution to storing keys
and certificates on a USB device without passwords or highlight
whether the risks of doing so are too high.

@@

Our test desktop machine was configured to accepted login
with a smartcard reader and an OpenPGP smartcard
\parencite{pietig2009functional}. The OpenPGP card integrates
with GnuPG to provide a public-private key pair that can
be used for encryption.

GnuPG has a gpg-agent service that allows these keys to
be used for SSH authentication as well, which can provide
a Single Sign-On experience betweeen logging in to a local
desktop and connecting out to remote servers.

For desktop login, we were able to use the ``Poldi'' PAM
module. The use of the smartcard requires a PIN and
the card has a lock-out after 3 incorrect tries. This is
behaviour users can recognise from their bank smartcards
and the lock-out feature negates the guessability of
even a weak 4 to 6 digit PIN.

The use of a PIN does not eliminate the ``something you
know'' factor, but a short PIN is far less of a burden
than the requirement for users to have strong, random
passwords. This could be a reasonable compromise for an
enterprise given some further usability research.

\section{YubiKey}
